On the specific recursive structure of iterated parametricity

### 2025-11-10 paper/paper.tex
paper: fix compile errors, send Hugo PDF

\begin{align*}
  \depstype[n][0,..,p-1] \defeq
  \left(\begin{array}{lll}
            \framep[n][0,...,p-1] &:&\frametype[n][0,...,p-1] \\
            \painting[n][0,...,p-1] &:&\paintingtype[n][0,...,p-1]
            (\framep[n][0,...,p-1])                         \\
            \restrf[n][0,...,p-1]  &:&\restrftype[n][0,...,p-1]
            \left(\begin{array}{l}
                \framep[n][0,...,p-1]   \\
                \painting[n][0,...,p-1] \\
              \end{array}\right)                   \\
          \end{array}\right)
\end{align*}
so that the type of $\framep[n][0,...,p]$ can be shortened into $\depstype[n][0,..,p-1] \rightarrow \frametype[n][0,...,p]$ and similarly for $\layertype[n][p]$ and $\layer[n][p]$.
Later, we shall write $\deps.\framep$, $\deps.\painting$, $\deps.\restrf$ for the corresponding projections of a $\deps$. By abuse of notation, we also write $\unitpoint$ for the only inhabitant of $\depstype[n][0,..,0-1]$.

@@

Let us introduce another abbreviation to combine a full $\deps[n][0,n]$ with a family $E:\framep[n-1][n-1]\rightarrow \U$:
\begin{align*}
  \fulldepstype[n] \defeq
  \left(\begin{array}{lll}
            \deps[n][0,...,n-1] &:&\depstype[n][0,...,n-1] \\
            E&:&\frame[n][n](\deps[n][0,...,n-1]) \rightarrow \U         \\
          \end{array}\right)
\end{align*}
TODO: say the convention on projection in the Introduction

@@

Thanks to $\restrptype[n][0,...,n]$, we can mutually define $\restrf[n][p]$ and $\cohftype[n][p]$ from ``full dependencies'', that is from $\frame[n][0,...,n]$, $\painting[n][0,...,n]$, $\restrf[n][0,...,n]$ and some $E:\frame[n+1][n+1](\, using abbreviations in passing for $\layer$ and $\layertype$:

@@

\begin{align*}
  \depscohstype[n][0,..,p-1] \defeq
  \left(\begin{array}{l}
            \fulldeps[n][0,...,n] &:&\fulldepstype[n][0,...,n]        \\
            \restrp[n][0,...p-1]&:& \restrptype[n][0,...,p-1] \\
            \cohf[n][0,...,p-1]&:& \cohftype[n][0,...,p-1]    \\
          \end{array}\right)
\end{align*}
so that the type of $\restrf[n][0,...,p]$ can be rephrased as $\Pi \depscohs[n][0,...,p-1]:\depscohstype[n][0,..,p-1]. \restrftype[n][0,...,p]      \left(\begin{array}{l}
      \depscohs.\fulldeps.\deps.\framep[n][0,...,p]   \\
      \depscohs.\fulldeps.\deps.\painting[n][0,...,p] \\
    \end{array}\right)$, and similarly for $\restrltype$ and $\restrl$. Later, we shall write $\depscohs.\deps$, $\depscohs.\restrp$, and $\depscohs.\cohf$ for the corresponding projections of an element $\depscohs$ in $\depscohstype$. By abuse of notation, we also write $\unitpoint$ for the only inhabitant of $\depscohstype[n][0,..,0-1]$.

@@

\begin{align*}
  \overline{\fulldeps}(\fulldepscohs:\fulldepscohstype) & :      & \fulldepstype \\
  \overline{\fulldeps}(\fulldepscohs)               & \defeq &
  \left(\begin{array}{l}
            \overline{\deps}(\fulldepscohs.\depscohs)\\
            \fulldepscohs.\depscohs.\fulldeps.E\\
          \end{array}\right)             \\
\end{align*}

We need also the notion of full dependencies at the level of coherence conditions, that combines a full $\depscohs[n][0,...,n]$ with a set $E:\framep[n][n](\overline{\deps}(\depscohs[n][0,...,n-1])) \rightarrow \U$:
\begin{align*}
  \fulldepscohstype[n] \defeq
  \left(\begin{array}{l}
            \depscohs[n][0,...,n-1] &:&\depscohstype[n][0,...,n-1]                         \\
            E&:&\framep[n+1][n+1](\overline{\deps}(\depscohs[n][0,...,n-1]))\rightarrow \U \\
          \end{array}\right)
\end{align*}
TODO: say the convention on projection in the Introduction

@@

\begin{align*}
  \depscohttype[n][0,..,p-1] \defeq
  \left(\begin{array}{lll}
            \fulldepscohs[n] &:&\fulldepscohstype[n]                        \\
            \cohp[n][0,...p-1]&:& \cohptype[n][0,...,p-1](\fulldepscohs[n]) \\
            \coht[n][0,...,p-1]&:& \cohttype[n][0,...,p-1]
            \left(\begin{array}{l}
                \fulldepscohs[n]    \\
                \cohp[n][0,...,p-1] \\
              \end{array}\right)                                     \\
          \end{array}\right)
\end{align*}

It is also convenient to build $\depscohs$ from $\depscohtt$ at the previous level:

\begin{align*}
  \overline{\depscohs}(\depscoht:\depscohttype) & :      & \depscohstype \\
  \overline{\depscohs}(\depscoht)               & \defeq &
  \left(\begin{array}{l}
            \overline{\deps}[n][p](\depscoht.\fulldepscohs.\depcohs[n][0,...,p])   \\
            \restrpainting[n][p](\depscoht.\fulldepscohs[n][0,...,n]) \\
            \cohf[n][p](\depscoht[n][0,...,p])         \\
          \end{array}\right)             \\
\end{align*}

\begin{align*}
  \overline{\fulldepscohs}(\fulldepscoht:\fulldepscohstype) & :      & \fulldepscohstype \\
  \overline{\fulldepscohs}(\fulldepscoht)               & \defeq &
  \left(\begin{array}{l}
            \overline{\depscohs}(\fulldepscoht.\depscoht)\\
            \fulldepscoht.\depscoht.\fulldepscohs.E\\
          \end{array}\right)             \\
\end{align*}

We need also the notion of full dependencies at the level of the coherence conditions of dimension 2, that combines a full $\depscoht[n][0,...,n]$ with a set $E:\framep[n][n](\overline{\depscohs}(\depscoht[n][0,...,n-1])) \rightarrow \U$:
\begin{align*}
  \fulldepscohttype[n] \defeq
  \left(\begin{array}{lll}
            \depscoht[n][0,...,n-1] &:&\depscohttype[n][0,...,n-1]    \\
            E&:&\framep[n+2][n+2](\overline{\depscohs}(\depscoht[n][0,...,n-1]))\rightarrow \U \\
          \end{array}\right)
\end{align*}

>>>>

\begin{align*}
  \depstype[n][0,..,p-1] \defeq
  \left(\begin{array}{lll}
            \framep[n][0,...,p-1]   & : & \frametype[n][0,...,p-1]    \\
            \painting[n][0,...,p-1] & : & \paintingtype[n][0,...,p-1]
            (\framep[n][0,...,p-1])                                   \\
            \restrf[n][0,...,p-1]   & : & \restrftype[n][0,...,p-1]
            \left(\begin{array}{l}
                \framep[n][0,...,p-1]   \\
                \painting[n][0,...,p-1] \\
              \end{array}\right)                             \\
          \end{array}\right)
\end{align*}
so that the type of $\framep[n][0,...,p]$ can be shortened into $\depstype[n][0,..,p-1] \rightarrow \frametype[n][0,...,p]$ and similarly for $\layertype[n][p]$ and $\layer[n][p]$.
Later, we shall write $\deps.\framep$, $\deps.\painting$, $\deps.\restrf$ for the corresponding projections of a $\deps$. By abuse of notation, we also write $\unitpoint$ for the only inhabitant of $\depstype[n][0,..,0-1]$.

@@

Let us introduce another abbreviation to combine a full $\deps[n][0,n]$ with a family $E:\framep[n-1][n-1]\rightarrow \U$:
\begin{align*}
  \fulldepstype[n] \defeq
  \left(\begin{array}{lll}
            \deps[n][0,...,n-1] & : & \depstype[n][0,...,n-1]                          \\
            E                   & : & \frame[n][n](\deps[n][0,...,n-1]) \rightarrow \U \\
          \end{array}\right)
\end{align*}
TODO: say the convention on projection in the Introduction

@@

Thanks to $\restrptype[n][0,...,n]$, we can mutually define $\restrf[n][p]$ and $\cohftype[n][p]$ from ``full dependencies'', that is from $\frame[n][0,...,n]$, $\painting[n][0,...,n]$, $\restrf[n][0,...,n]$ and some $E:\frame[n+1][n+1]$, using abbreviations in passing for $\layer$ and $\layertype$:

@@

\begin{align*}
  \depscohstype[n][0,..,p-1] \defeq
  \left(\begin{array}{lcl}
            \fulldeps[n][0,...,n] & : & \fulldepstype[n][0,...,n] \\
            \restrp[n][0,...p-1]  & : & \restrptype[n][0,...,p-1] \\
            \cohf[n][0,...,p-1]   & : & \cohftype[n][0,...,p-1]   \\
          \end{array}\right)
\end{align*}
so that the type of $\restrf[n][0,...,p]$ can be rephrased as $\Pi \depscohs[n][0,...,p-1]:\depscohstype[n][0,..,p-1]. \restrftype[n][0,...,p]      \left(\begin{array}{l}
      \depscohs.\fulldeps.\deps.\framep[n][0,...,p]   \\
      \depscohs.\fulldeps.\deps.\painting[n][0,...,p] \\
    \end{array}\right)$, and similarly for $\restrltype$ and $\restrl$. Later, we shall write $\depscohs.\deps$, $\depscohs.\restrp$, and $\depscohs.\cohf$ for the corresponding projections of an element $\depscohs$ in $\depscohstype$. By abuse of notation, we also write $\unitpoint$ for the only inhabitant of $\depscohstype[n][0,..,0-1]$.

@@

\begin{align*}
  \overline{\fulldeps}(\fulldepscohs:\fulldepscohstype) & :      & \fulldepstype \\
  \overline{\fulldeps}(\fulldepscohs)                   & \defeq &
  \left(\begin{array}{l}
            \overline{\deps}(\fulldepscohs.\depscohs) \\
            \fulldepscohs.\depscohs.\fulldeps.E       \\
          \end{array}\right)                                \\
\end{align*}

We need also the notion of full dependencies at the level of coherence conditions, that combines a full $\depscohs[n][0,...,n]$ with a set $E:\framep[n][n](\overline{\deps}(\depscohs[n][0,...,n-1])) \rightarrow \U$:
\begin{align*}
  \fulldepscohstype[n] \defeq
  \left(\begin{array}{lcl}
            \depscohs[n][0,...,n-1] & : & \depscohstype[n][0,...,n-1]                                                \\
            E                       & : & \framep[n+1][n+1](\overline{\deps}(\depscohs[n][0,...,n-1]))\rightarrow \U \\
          \end{array}\right)
\end{align*}
TODO: say the convention on projection in the Introduction

@@

\begin{align*}
  \depscohttype[n][0,..,p-1] \defeq
  \left(\begin{array}{lll}
            \fulldepscohs[n]    & : & \fulldepscohstype[n]                      \\
            \cohp[n][0,...p-1]  & : & \cohptype[n][0,...,p-1](\fulldepscohs[n]) \\
            \coht[n][0,...,p-1] & : & \cohttype[n][0,...,p-1]
            \left(\begin{array}{l}
                \fulldepscohs[n]    \\
                \cohp[n][0,...,p-1] \\
              \end{array}\right)                                           \\
          \end{array}\right)
\end{align*}

It is also convenient to build $\depscohs$ from $\depscoht$ at the previous level:

\begin{align*}
  \overline{\depscohs}(\depscoht:\depscohttype) & :      & \depscohstype      \\
  \overline{\depscohs}(\depscoht)               & \defeq &
  \left(\begin{array}{l}
            \overline{\deps}[n][p](\depscoht.\fulldepscohs.\depscohs[n][0,...,p]) \\
            \restrp[n][p](\depscoht.\fulldepscohs[n][0,...,n])                    \\
            \cohf[n][p](\depscoht[n][0,...,p])                                    \\
          \end{array}\right) \\
\end{align*}

\begin{align*}
  \overline{\fulldepscohs}(\fulldepscoht:\fulldepscohstype) & :      & \fulldepscohstype \\
  \overline{\fulldepscohs}(\fulldepscoht)                   & \defeq &
  \left(\begin{array}{l}
            \overline{\depscohs}(\fulldepscoht.\depscoht) \\
            \fulldepscoht.\depscoht.\fulldepscohs.E       \\
          \end{array}\right)                                    \\
\end{align*}

We need also the notion of full dependencies at the level of the coherence conditions of dimension 2, that combines a full $\depscoht[n][0,...,n]$ with a set $E:\framep[n][n](\overline{\depscohs}(\depscoht[n][0,...,n-1])) \rightarrow \U$:
\begin{align*}
  \fulldepscohttype[n] \defeq
  \left(\begin{array}{lll}
            \depscoht[n][0,...,n-1] & : & \depscohttype[n][0,...,n-1]                                                    \\
            E                       & : & \framep[n+2][n+2](\overline{\depscohs}(\depscoht[n][0,...,n-1]))\rightarrow \U \\
          \end{array}\right)
\end{align*}
